--
title: "DSC 200 Lab Project -Term 2232"
date: "`r Sys.Date()`"
output: github_document
---

**Student Name:salwa alshmry**

**Student ID:<2221001664**


**Deadline:** 23:59 on Sunday, 19 May 2024

**Total Points:** 20


## Loading Packages

```{r load-packages, message = FALSE}

library(tidyverse)
library(openintro)
library(ggrepel)
```


## Tasks

`1. (2 points)
```{r}
# Count the number of pets
num_pets <- nrow(seattlepets)
num_pets
```
Write your narrative here




`2. (2 points)

```{r}

```
Write your narrative below



`3. (2 points)
```{r}

```
Write your narrative here


`4. (2 points)
```{r}

```
Write your narrative here


`5. (2 points)
```{r}

```
Write your narrative here




`6. (2 points)
```{r}

```
Write your narrative here





`7. (2 points)
```{r}

```
Write your narrative here



`8. (2 points)
```{r}

```
Write your narrative here




`9. (2 points)
```{r}
top_10_names <- seattlepets %>% 
filter(animal_name %in% c( "Lucy"  , "Charlie" , "Luna" , "Bella" , "Max"    , 
                           "Daisy" , "Molly"   , "Jack" , "Lily"  , "Stella" ))
top_10_names
```
`a.	What does the above code chunk do?



`b. Plot the counts of the pet names (animal_name) in top_10_names
```{r}

```


`10.  (2 points)

`The below code plots the proportion of dogs with a given name versus the proportion of cats with the same name. The 20 most common cat and dog names are displayed. The diagonal line on the plot is the x = y line; if a name appeared on this line, the name’s popularity would be exactly the same for dogs and cats.

```{r eval = TRUE, echo = FALSE, fig.fullwidth = TRUE}
# data prep
name_props <- seattlepets %>% 
  filter(
    !is.na(animal_name),
    species %in% c("Dog", "Cat")
    ) %>%
  group_by(species) %>% 
  count(animal_name, sort = TRUE) %>%
  mutate(prop = n / sum(n))
    
cat_name_props <- name_props %>%
  filter(species == "Cat") %>%
  rename(cat_prop = prop) %>%
  slice(1:30)

dog_name_props <- name_props %>%
  filter(species == "Dog") %>%
  rename(dog_prop = prop) %>%
  slice(1:30)
    
comb_name_props <- inner_join(cat_name_props, dog_name_props, 
                              by = "animal_name") %>%
  ungroup() %>%
  select(animal_name, cat_prop, dog_prop)
    
# create viz
ggplot(comb_name_props, aes(x = cat_prop, y = dog_prop)) +
  geom_abline(intercept = 0, 
              color = COL["lgray","full"], 
              alpha = 0.8, 
              size = 1.5) +
  geom_text_repel(aes(label = animal_name), 
                  segment.color = COL["gray","full"], 
                  seed = 291252, max.iter = 10000) +
  geom_point(color = COL["blue","full"], alpha = 0.8) +
  theme_minimal() +
  labs(x = "Proportion of cats", y = "Proportion of dogs") +
  xlim(0.002, 0.01) +
  ylim(0.002, 0.01) +
  ggimage::geom_emoji(
      image = "1f436", aes(x = 0.003, y = 0.009), size = 0.1
      ) +
  ggimage::geom_emoji(
      image = "1f431", aes(x = 0.009, y = 0.003), size = 0.1
      )
```

`What names are more common for cats than dogs? The ones above the line or the ones below the line?

`Answer here 
The names that are more common for cats than dogs are the ones above the diagonal line (x = y) on the plot.


`Is the relationship between the two variables (proportion of cats with a given name and proportion of dogs with a given name) positive or negative? What does this mean in context of the data?

`Answer here 
The relationship between the proportion of cats with a given name and the proportion of dogs with the same name is negative. In the context of the data, this negative relationship means that as the proportion of cats with a given name increases, the proportion of dogs with the same name tends to decrease, and vice versa. This suggests an inverse relationship in the popularity of names between cats and dogs in the dataset.
