--
title: "DSC 200 Lab Project -Term 2232"
date: "`r Sys.Date()`"
output: github_document
---

**Student Name:salwa alshmry**

**Student ID:<2221001664**


**Deadline:** 23:59 on Sunday, 19 May 2024

**Total Points:** 20


## Loading Packages

```{r load-packages, message = FALSE}

library(tidyverse)
library(openintro)
library(ggrepel)
```


## Tasks
`1. (2 points)
```{r}
num_pets <- nrow(seattlepets)
num_pets
```
Counting Pets:
num_pets <- nrow(seattlepets): Counts the rows in the dataset seattlepets, storing the count in num_pets.
Displaying Count:
num_pets: Displays the total count of pets stored in num_pets.
Brief Summary:
This code counts the total number of pets in the dataset seattlepets and displays the count. Each row represents one pet in the dataset.
`2. (2 points)
```{r}
# Count the number of variables num_variables <- ncol(seattlepets)
num_variables
```
Counting Variables:
num_variables <- ncol(seattlepets): Calculates the number of variables (columns) in the seattlepets dataset.
The ncol() function returns the number of variables in the data frame passed to it, which in this case is seattlepets.
The resulting number is stored in the variable num_variables.
Displaying Count:
num_variables: Displays the total count of variables stored in num_variables.
Brief Summary:
This code counts the total number of variables in the seattlepets dataset and displays the count. Each column represents one variable in the dataset.
`3. (2 points)
```{r}
#Count the frequency of each species species_count <- table(seattlepets$species)
species_count
```
Counting Species Frequency:
species_count <- table(seattlepets$species): Calculates the frequency of each species in the seattlepets dataset.
The table() function creates a frequency table of the specified variable, which in this case is species extracted from the seattlepets dataset.
It counts the occurrences of each unique value in the species column and returns a table summarizing the counts.
The resulting frequency table is stored in the variable species_count.
Displaying Frequency Table:
species_count: Displays the frequency table showing the count of each species.
Brief Summary:
This code calculates the frequency of each species recorded in the seattlepets dataset and displays the frequency table. It provides insights into the distribution of different species among the registered pets.
`4. (2 points)
```{r}
# Count the frequency of each pet name
pet_name_frequency <- table(seattlepets$animal_name)
# Sort the frequency table in descending order
sorted_pet_names <- sort(pet_name_frequency, decreasing = TRUE)
# Display the top ten most common  names
head(sorted_pet_names, 10)
```
Counting Pet Name Frequency:
pet_name_frequency <- table(seattlepets$animal_name): Calculates the frequency of each pet name in the seattlepets dataset.
The table() function creates a frequency table of the specified variable, which in this case is animal_name extracted from the seattlepets dataset.
It counts the occurrences of each unique pet name in the animal_name column and returns a table summarizing the counts.
The resulting frequency table is stored in the variable pet_name_frequency.
Sorting the Frequency Table:
sorted_pet_names <- sort(pet_name_frequency, decreasing = TRUE): Sorts the frequency table pet_name_frequency in descending order based on the count of each pet name.
The sort() function arranges the elements of the frequency table in descending order, with the most common pet names appearing first.
The parameter decreasing = TRUE specifies that the sorting should be done in descending order.
Displaying the Top Ten Most Common Pet Names:
head(sorted_pet_names, 10): Displays the top ten most common pet names from the sorted frequency table sorted_pet_names.
The head() function extracts the first ten rows from the frequency table, representing the top ten most common pet names based on their frequency of occurrence.
Brief Summary:
This code calculates the frequency of each pet name recorded in the seattlepets dataset, sorts them in descending order based on their frequency, and displays the top ten most common pet names. It provides insights into the popularity of different pet names among registered pets.
`5. (2 points)
```{r}
# Filter records for species Pig
pig_records <- filter(seattlepets, species == "Pig")
# Arrange the filtered records by pet names (animal_name)
sorted_pig_records <- arrange(pig_records, animal_name)
# Display the sorted pig records
sorted_pig_records
```
Filtering Records for Species Pig:
pig_records <- filter(seattlepets, species == "Pig"): Filters the records in the seattlepets dataset to only include those where the species is “Pig”.
The filter() function selects rows from the dataset where the condition species == "Pig" is met.
It creates a subset of the dataset containing only records for pigs.
The filtered records are stored in the variable pig_records.
Arranging Filtered Records by Pet Names:
sorted_pig_records <- arrange(pig_records, animal_name): Arranges the filtered pig records by pet names (animal_name) in ascending order.
The arrange() function sorts the rows of the dataset based on the specified column, which in this case is animal_name.
The records are sorted alphabetically by pet names.
The sorted records are stored in the variable sorted_pig_records.
Displaying the Sorted Pig Records:
sorted_pig_records: Displays the filtered pig records sorted by pet names.
This command shows the subset of data containing information about pigs, arranged alphabetically by their pet names.
Brief Summary:
This code filters the records in the seattlepets dataset to include only those for the species “Pig”, arranges these filtered records alphabetically by pet names, and then displays the sorted pig records. It provides a sorted list of pig records, making it easier to analyze and interpret the data related to pigs.
`6. (2 points)
```{r}
# Select only pet name (animal_name) and primary breed for species Goat goat_records <- select(filter(seattlepets, species == "Goat"), animal_name, primary_breed)
# Arrange the selected records by pet names (animal_name)sorted_goat_records <- arrange(goat_records, animal_name)
# Display the sorted goat records
sorted_goat_records
```
Selecting Pet Name and Primary Breed for Species Goat:
goat_records <- select(filter(seattlepets, species == "Goat"), animal_name, primary_breed): Selects only the columns animal_name and primary_breed from the seattlepets dataset for records where the species is “Goat”.
The filter() function is used to subset the dataset to include only records where the species is “Goat”.
The select() function then selects the specified columns (animal_name and primary_breed) from the filtered dataset.
The selected records are stored in the variable goat_records.
Arranging Selected Records by Pet Names:
sorted_goat_records <- arrange(goat_records, animal_name): Arranges the selected goat records by pet names (animal_name) in ascending order.
The arrange() function sorts the rows of the dataset based on the specified column, which in this case is animal_name.
The records are sorted alphabetically by pet names.
The sorted records are stored in the variable sorted_goat_records.
Displaying the Sorted Goat Records:
sorted_goat_records: Displays the selected goat records sorted by pet names.
This command shows the subset of data containing information about goats, with columns for pet names and primary breeds, arranged alphabetically by pet names.
Brief Summary:
This code selects and extracts specific columns (pet name and primary breed) from the seattlepets dataset for records where the species is “Goat”, arranges these selected records alphabetically by pet names, and then displays the sorted goat records. It provides a focused view of goat-related data, including pet names and primary breeds, facilitating analysis and interpretation.
`7. (2 points)
```{r}
# Concatenate animal_name and species into a single column named pet
seattlepets <- mutate(seattlepets, pet = paste(animal_name, species, sep = " - "))
# Select license_number and pet, then arrange by pet
sorted_records <- select(seattlepets, license_number, pet) %>%
                  arrange(pet)
# Display the sorted records
sorted_records
```
Loading the Required Library:
library(dplyr): Loads the dplyr library, which provides functions for data manipulation.
Concatenating Columns into a Single Column:
seattlepets <- mutate(seattlepets, pet = paste(animal_name, species, sep = " - ")): Creates a new column named “pet” in the seattlepets dataset by concatenating the columns animal_name and species with a separator (“-”).
The mutate() function is used to add new variables or modify existing variables in the dataset.
The paste() function concatenates the values of animal_name and species, with ” - ” as the separator.
The resulting concatenated values are stored in the new column “pet”.
Selecting and Arranging Columns:
select(seattlepets, license_number, pet) %>% arrange(pet): Selects the columns license_number and pet from the modified seattlepets dataset and arranges the rows by the column “pet” in ascending order.
The %>% operator (pipe) is used to perform multiple operations sequentially.
The arrange() function sorts the rows of the dataset based on the specified column, which in this case is “pet”.
The selected and sorted records are stored in the variable sorted_records.
Displaying the Sorted Records:
sorted_records: Displays the selected and sorted records.
This command shows the subset of data containing information about license numbers and pets (concatenated pet names and species), arranged alphabetically by the “pet” column.
Brief Summary:
This code concatenates the columns animal_name and species into a single column named “pet” in the seattlepets dataset, selects and arranges the columns for license numbers and pets, and then displays the sorted records. It provides a consolidated view of pet information, facilitating further analysis and interpretation.
`8. (2 points)
```{r}
library(ggplot2)
# Create a bar plot of species counts
species_plot <- ggplot(seattlepets, aes(x = species)) +
                geom_bar() +
                labs(title = "Counts of Species", x = "Species", y = "Count") +
                theme(axis.text.x = element_text(angle = 45, hjust = 1))
# Display the plot
print(species_plot)
```


Loading the Required Library:
library(ggplot2): Loads the ggplot2 library, which provides functions for creating data visualizations.
Creating a Bar Plot of Species Counts:
species_plot <- ggplot(seattlepets, aes(x = species)) + geom_bar() + labs(title = "Counts of Species", x = "Species", y = "Count") + theme(axis.text.x = element_text(angle = 45, hjust = 1)):
ggplot() initializes a new ggplot object.
seattlepets specifies the dataset to be used.
aes(x = species) maps the x-axis to the “species” column in the dataset.
geom_bar() creates a bar plot where the height of each bar represents the count of occurrences of each species.
labs() sets the title and axis labels for the plot.
theme(axis.text.x = element_text(angle = 45, hjust = 1)) adjusts the appearance of the x-axis text by rotating it 45 degrees and aligning it to the right.
Displaying the Plot:
print(species_plot): Displays the created bar plot.
This command prints the ggplot object species_plot, generating the bar plot visualization of species counts with the specified title, axis labels, and theme settings.
Brief Summary:
This code utilizes ggplot2 to create a bar plot of species counts based on the seattlepets dataset. The x-axis represents different species, and the height of each bar corresponds to the count of occurrences of that species. The plot is then displayed with appropriate title, axis labels, and theme settings.
`9. (2 points)
```{r}
top_10_names <- seattlepets %>% 
filter(animal_name %in% c( "Lucy"  , "Charlie" , "Luna" , "Bella" , "Max"    , 
                           "Daisy" , "Molly"   , "Jack" , "Lily"  , "Stella" ))
top_10_names
```
`a.	What does the above code chunk do?
# Create a subset of data for the top 10 most common pet names
```{r}
top_10_names <- c("Lucy", "Charlie", "Luna", "Bella", "Max", "Cooper",
"Daisy", "Buddy", "Molly", "Stella")
top_10_subset <- filter(seattlepets, animal_name %in% top_10_names)
```
`b. Plot the counts of the pet names (animal_name) in top_10_names
# Plot the counts of the pet names segmented by species
```{r}
pet_names_plot <- ggplot(top_10_subset, aes(x = animal_name, fill =
species)) +
geom_bar(position = "dodge") +
labs(title = "Counts of Top 10 Pet Names Segmented by
Species", x = "Pet Name", y = "Count") +
theme(axis.text.x = element_text(angle = 45, hjust =
1))
# Display the plot
print(pet_names_plot)
```
`10.  (2 points)
`The below code plots the proportion of dogs with a given name versus the proportion of cats with the same name. The 20 most common cat and dog names are displayed. The diagonal line on the plot is the x = y line; if a name appeared on this line, the name’s popularity would be exactly the same for dogs and cats.

```{r eval = TRUE, echo = FALSE, fig.fullwidth = TRUE}
# data prep
name_props <- seattlepets %>% 
  filter(
    !is.na(animal_name),
    species %in% c("Dog", "Cat")
    ) %>%
  group_by(species) %>% 
  count(animal_name, sort = TRUE) %>%
  mutate(prop = n / sum(n))
    
cat_name_props <- name_props %>%
  filter(species == "Cat") %>%
  rename(cat_prop = prop) %>%
  slice(1:30)

dog_name_props <- name_props %>%
  filter(species == "Dog") %>%
  rename(dog_prop = prop) %>%
  slice(1:30)
    
comb_name_props <- inner_join(cat_name_props, dog_name_props, 
                              by = "animal_name") %>%
  ungroup() %>%
  select(animal_name, cat_prop, dog_prop)
    
# create viz
ggplot(comb_name_props, aes(x = cat_prop, y = dog_prop)) +
  geom_abline(intercept = 0, 
              color = COL["lgray","full"], 
              alpha = 0.8, 
              size = 1.5) +
  geom_text_repel(aes(label = animal_name), 
                  segment.color = COL["gray","full"], 
                  seed = 291252, max.iter = 10000) +
  geom_point(color = COL["blue","full"], alpha = 0.8) +
  theme_minimal() +
  labs(x = "Proportion of cats", y = "Proportion of dogs") +
  xlim(0.002, 0.01) +
  ylim(0.002, 0.01) +
  ggimage::geom_emoji(
      image = "1f436", aes(x = 0.003, y = 0.009), size = 0.1
      ) +
  ggimage::geom_emoji(
      image = "1f431", aes(x = 0.009, y = 0.003), size = 0.1
      )
```

`What names are more common for cats than dogs? The ones above the line or the ones below the line?

`Answer here 
The names that are more common for cats than dogs are the ones above the diagonal line (x = y) on the plot.


`Is the relationship between the two variables (proportion of cats with a given name and proportion of dogs with a given name) positive or negative? What does this mean in context of the data?

`Answer here 
The relationship between the proportion of cats with a given name and the proportion of dogs with the same name is negative. In the context of the data, this negative relationship means that as the proportion of cats with a given name increases, the proportion of dogs with the same name tends to decrease, and vice versa. This suggests an inverse relationship in the popularity of names between cats and dogs in the dataset.
